---
title: "R_Unmarked"
author: "Jeff Grayum"
date: "2/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries/packages.
```{r}
library(tidyverse)
library(scales)
library(ggthemes)
#install.packages("unmarked")
library(unmarked)
#install.packages("auk")
library(auk)
library(lubridate)
#install.packages("raster")
library(raster)
#installed.packages("dgGridr")
library(dgGridr)
#install.packages("ebirdst")
library(ebirdst)
#install.packages("fields")
library(fields)
#install.packages("AICcmodavg")
library(AICcmodavg)
library(sf)
#install.packages("rgdal")
library(rgdal)
```

-Often we won't know how many any individuals are occupying a study area.  We only know how many have been detected at each survey point.  Other individuals? Overlap between detections?
-Occupancy analysis allows us to use these detections to get an idea of actual abundance.
-In the video example, Bushrats are detected at 0.5 (50%) of sites.  0.5 is the "Naive Occupancy"

p = detection probability (CHANCE OF DETECTING AN ANIMAL IF THEY OCCUR THERE)
psi = occupancy.

-Occupancy analysis takes into acct our imperfect detection prob, and studies the daily pattern of detection and non-detection at a site.  Also allows us to calculate Confidence Intervals.


Example analysis using an eBird data set (I hope!!)
```{r}
#First, we must resolve some namespace conflicts between packages.
select <- dplyr::select
projection <- raster::projection
map <- purrr::map

#Next, we set a random number seed.  This should make our results more repeatable!
set.seed(1)

#We then setup an output director for saved results.
if(!dir.exists("output")) {
  dir.create("output")
}

#Import data from eBird. Note that occupancy modeling requires an integer response
ebird <- read_csv("ebd_woothr_june_bcr27_zf.csv") %>%
  mutate(year = year(observation_date),
                     species_observed = as.integer(species_observed))

#Deviating from lesson to get a feel for the data
ebird %>%
  count(scientific_name, sort = TRUE)

ebird %>%
  count(observer_id, sort = TRUE)

ebird %>%
  count(state_code, sort = TRUE)

#Ok. So we are modeling data from a single bird species over 11 states, with data from many observers.

#Loading modis land cover covariates
habitat <- read_csv("pland-elev_location-year.csv") %>%
  mutate(year = as.integer(year))

habitat %>%
  view()

#Joining eBird and modis data.
ebird_habitat <- inner_join(ebird, habitat, by = c("locality_id", "year"))

#Loading prediction surface
pred_surface <- read_csv("pland-elev_prediction-surface.csv")

#Latest year of landcover data.
max_lc_year <- pred_surface$year[1]
r <- raster("prediction-surface.tif")

#Let's load our gis data for making maps.
map_proj <- st_crs(102003)

ne_land <- read_sf("gis-data.gpkg", "ne_land") %>%
  st_transform(crs = map_proj) %>%
  st_geometry()
#Running into issues here with crs not being found... but need to eat breakfast.

```


Haven't had any luck troubleshooting.  Will try working through a different example, hopefully can figure out why crs isnt found after I learn more.

Link to example: https://doi90.github.io/lodestar/fitting-occupancy-models-with-unmarked.html
```{r}
#Changes in occupancy are linked to changes in population size as well as pop'n characteristics.
#IMPERFECT DETECTION -> why these occupancy models are so important.
#State process: defined by the true occupancy status (site occupied, or it's not..)
#Observation process: was the species detected?
#covariates can be incorporated into the model using  a logit-link function.
#Assumptions: Site is closed to changes in occupancy during season, prob of occupance is constant across all sites or VARATION IS MODELLED BY COVARIATES, prob of detection is constant across all sites, or variation is modeled by covariates!!, detection of a species and its detection histories at each site are independent.

#This guide will only cover occu() for standand occupancy models.

#This example will use a dataset built into unmarked::, which is handy.

data <- read.csv(system.file("csv", "widewt.csv", package = "unmarked"))

data %>%
  view()


#unmarkedFrame formats data to a specific model fitting function.  A special class of object used in this package for organizing our data.
#Unmarked helps users fit models accounting for the imperfect detection of unmarked individuals.
#Can also do standard occupancy models, distance sampling, and colonization/extinction models.

```

-unmarkedFrameOccu requires: 2 dataframes, a list of dataframes.
--> y: dataframe of presense absence records.  Rows are sites, columns are repeat visits.
--> siteCovs: a datafrome of the site-level covariates.  Things that DON"T change between visits to a particular site(elevation, annual rainfall, distance from road, etc.).  One column per covariate.
--> obsCovs:  A list of dataframes for the observation level covariates.  Each covariate is it's own dataframe, rows are sites, and columns are repeat visits..  Might be environmental conditions like weather, or methodological variables like survey method (camera trap, pitfall trap, spotlighting, etc).  Think of having datafram for lunar illumination, with rows as different sites, columns as repeat visits.  Another dataframe frame for canopy cover, with rows as sites and columns as repeat visits.  Another dataframe for temperature, with rows as sites and columns as complete visits.  All these various dataframes would be in a list of obsCovs.
```{r}

```

-=