---
title: "R_Unmarked"
author: "Jeff Grayum"
date: "2/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries/packages.
```{r}
library(tidyverse)
library(scales)
library(ggthemes)
#install.packages("unmarked")
library(unmarked)
#install.packages("auk")
library(auk)
library(lubridate)
#install.packages("raster")
library(raster)
#installed.packages("dgGridr")
library(dgGridr)
#install.packages("ebirdst")
library(ebirdst)
#install.packages("fields")
library(fields)
#install.packages("AICcmodavg")
library(AICcmodavg)
library(sf)
#install.packages("rgdal")
library(rgdal)
```

-Often we won't know how many any individuals are occupying a study area.  We only know how many have been detected at each survey point.  Other individuals? Overlap between detections?
-Occupancy analysis allows us to use these detections to get an idea of actual abundance.
-In the video example, Bushrats are detected at 0.5 (50%) of sites.  0.5 is the "Naive Occupancy"

p = detection probability (CHANCE OF DETECTING AN ANIMAL IF THEY OCCUR THERE)
psi = occupancy.

-Occupancy analysis takes into acct our imperfect detection prob, and studies the daily pattern of detection and non-detection at a site.  Also allows us to calculate Confidence Intervals.


Example analysis using an eBird data set (I hope!!)
```{r}
#First, we must resolve some namespace conflicts between packages.
select <- dplyr::select()
projection <- raster::projection()

#Next, we set a random number seed.  This should make our results more repeatable!
set.seed(1)

#We then setup an output director for saved results.
if(!dir.exists("output")) {
  dir.create("output")
}

#Import data from eBird. Note that occupancy modeling requires an integer response
ebird <- read_csv("ebd_woothr_june_bcr27_zf.csv") %>%
  mutate(year = year(observation_date),
                     species_observed = as.integer(species_observed))

#Deviating from lesson to get a feel for the data
ebird %>%
  count(scientific_name, sort = TRUE)

ebird %>%
  count(observer_id, sort = TRUE)

ebird %>%
  count(state_code, sort = TRUE)

#Ok. So we are modeling data from a single bird species over 11 states, with data from many observers.

#Loading modis land cover covariates
habitat <- read_csv("pland-elev_location-year.csv") %>%
  mutate(year = as.integer(year))

habitat %>%
  view()

#Joining eBird and modis data.
ebird_habitat <- inner_join(ebird, habitat, by = c("locality_id", "year"))

#Loading prediction surface
pred_surface <- read_csv("pland-elev_prediction-surface.csv")

#Latest year of landcover data.
max_lc_year <- pred_surface$year[1]
r <- raster("prediction-surface.tif")

#Let's load our gis data for making maps.
map_proj <- st_crs(102003)

ne_land <- read_sf("gis-data.gpkg", "ne_land") %>%
  st_transform(crs = map_proj) %>%
  st_geometry()
#Running into issues here with crs not being found... but need to eat breakfast.

```


-=